Rule Synthesis for Rule-based Intrusion Detectors

### 2020-05-18 paper/paper.tex

%\vspace{-1ex}
\begin{table}[t!]
  \footnotesize
  \caption{\label{table:rules}Example of non-documentation options of Suricata.}
  \vspace{-2ex}
  \centering
  \begin{tabular}{clp{6cm}}
    \toprule
    \multicolumn{1}{c}{\#} & \multicolumn{1}{c}{Name} &  \multicolumn{1}{c}{Description}\\
    \midrule     
    1 & dsize & matches on the size of the packet payload\\
    2 & itype &  matches on a specific ICMP type\\
    3 & icode & matches on a specific ICMP code\\
    4 & icmp\_seq  & checks a ICMP sequence number\\
    5 & icmp\_id & matches on specific ICMP id-values\\
    6 & window & checks the size of the TCP window\\
    7 & flags & matches on TCP flags\\
    8 & fragbits & checks if the frag. or reserved bits are set in the IP header\\
    9 & threshold & minimum threshold for a rule before it generates alerts\\
    10 & content & checks if argument is present on the payload\\
    \bottomrule
  \end{tabular}
  \vspace{-3ex}  
\end{table}

Table~\ref{table:rules} shows a sample of non-documentation options \rui{what 
is the meaning of non-documentation in this context?} \Luc{documentation options
are options that gives information about the rule itself and are not used for 
matching on packet data. So, non-documentation options are the ones that really
matters for detection.} of Suricata. Column ``\#'' shows the option number, column 
``Name'' shows the name of the option, and column ``Description'' summarizes
purpose. The option \CodeIn{content} deserves attention. It is common
in rules to capture http attacks and http rules are the most common in
the Emerging Threats Open Ruleset~\cite{emerging-threats-open}. The
content option takes a sequence of bytes as argument and checks if
that sequence is in the payload of the message.  For example, the
option \CodeIn{content:''SELECT''} checks if the string ``SELECT'' is
present in the payload.  Detailed description of the Suricata rules
can be found elsewhere~\cite{suri-rule-format}. Suricata supports a
total of \totoptions\ distinct options covering \numproto\ distinct
network protocols.

@@

In the following, we briefly show how \tname{} synthesizes rules for
this attack. In short, \tname{} takes one or more examples of an
attack---in the form of network messages---and produces rules to
capture that attack. Considering the rule for the Ping Scan attack,
whose example was reproduced with NMAP, \tname{} leaves the action of
the corresponding rule constant (as \CodeIn{alert}) and infers the
header by checking the protocol of the input messages. For example, in
this case, \tname{} sets the header to \CodeIn{icmp any any -> any
  any}, indicating that the rule applies to \CodeIn{icmp} messages and
the most general flow description, \ie{}, it instructs the \nids\ to
analyze the icmp traffic flowing from/to any address and port.  The
rationale for this decision is that the choice of these parameters are
domain-specific and subject to change by the system administrator.

>>>>

%\vspace{-1ex}
\begin{table}[t!]
  \footnotesize
  \caption{\label{table:rules}Example of capture-relevant options of Suricata.}
  \vspace{-2ex}
  \centering
  \begin{tabular}{clp{6cm}}
    \toprule
    \multicolumn{1}{c}{\#} & \multicolumn{1}{c}{Name} &  \multicolumn{1}{c}{Description}\\
    \midrule     
    1 & dsize & matches on the size of the packet payload\\
    2 & itype &  matches on a specific ICMP type\\
    3 & icode & matches on a specific ICMP code\\
    4 & icmp\_seq  & checks a ICMP sequence number\\
    5 & icmp\_id & matches on specific ICMP id-values\\
    6 & window & checks the size of the TCP window\\
    7 & flags & matches on TCP flags\\
    8 & fragbits & checks if the frag. or reserved bits are set in the IP header\\
    9 & threshold & minimum threshold for a rule before it generates alerts\\
    10 & content & checks if argument is present on the payload\\
    \bottomrule
  \end{tabular}
  \vspace{-3ex}  
\end{table}

Table~\ref{table:rules} shows a sample of capture-relevant options of
Suricata. Column ``\#'' shows the option number, column ``Name'' shows
the name of the option, and column ``Description'' summarizes
purpose. The option \CodeIn{content} deserves special attention. It is
common in rules to capture http attacks and http rules are the most
common in the Emerging Threats Open
Ruleset~\cite{emerging-threats-open}. The content option takes a
sequence of bytes as argument and checks if that sequence appears in
the payload of the message.  For example, the option
\CodeIn{content:''SELECT''} checks if the string ``SELECT'' is present
in the payload.  Detailed description of the Suricata rules can be
found elsewhere~\cite{suri-rule-format}. Suricata supports a total of
\totoptions\ distinct options covering \numproto\ distinct network
protocols.

@@

%---in the form of network messages---
% one or more examples of an attack
%, indicating that the rule applies to \CodeIn{icmp}messages

In the following, we show how \tname{} synthesizes rules for this
attack. \tname{} takes on input benign and malicious traffic and
produces on output rules to capture only the malicious
traffic. Considering the Ping Scan attack, \tname{} replays the
attack---with NMAP, in this case---and monitors the traffic.  It
infers the header by checking the protocol of the input messages. In
this case, \tname{} sets the header to \CodeIn{icmp any any -> any
  any}. It uses the most general flow description instructing the
\nids\ to analyze the icmp traffic flowing from/to any address and
port.  The rationale is that the choice of these parameters are
domain-specific and subject to change by the sys admins. Likewise,
\tname{} leaves the action of the corresponding rule constant, as
\CodeIn{alert}.

\Mar{--------------------- I am here...}
